StarCraft Bots and Competitions

### 2015-12-16 ECGG_StarCraftCompetitions/ECGG_StarCraftCompetitions.tex
small fix to numbers + rewording sscai intro

In 2011 the University of Alberta hosted the competition, with organization by Michael Buro and
David Churchill\footnote{\url{https://skatgame.net/mburo/sc2011/}}. Due to a lack of entrants in tournament categories 1-3 in the 2010 competition, it was
decided that only the full game category would be played in the 2011 competition. Another important
change in the 2011 competition was the introduction of automated tournament-managing software running
StarCraft games simultaneously on 20 computers, allowing a total of 1170 games to be played in the five days that the tournament ran. 
This increase in games played also allowed the tournament
to switch to a round-robin format, eliminating the ``luck'' factor of the pairings inherent in bracket
style tournaments. The bot that achieved the highest win percentage over the course of the competition would
be determined the winner. Also, the competition became open-source, in an effort 
to prevent possible cheating and to promote healthy competition in future tournaments by giving
newcomers and easier entry point by basing their design off of previous bots.

@@

The University of Alberta also hosted the 2012 competition, with the major difference from the 2011
competition being the addition of file reading and writing for learning. Bots could now write information to disk during a
match, and then read the information during other matches, allowing them to adjust strategies based
on previous results. 6 of the 10 entrants used this feature to aid in strategy selection, including the
top 4 finishers. More improvements to the tournament environment also meant that a total of 4240 games
could now be played in the same time period. Results are shown in Table \ref{tab:aiide2012}.

@@

A total of 8 bots competed in the 2013 AIIDE competition with many of the same names from the 2012 competition, and some key updates were made to existing bots for 2013 which shook up the results from the previous years. UAlbertaBot took 1st place with a dominant 84.49\% win rate, with Skynet in 2nd with 72.77\%. The major addition to UAlbertaBot was a combat simulation package called SparCraft. UAlbertaBot was re-engineered so that it now always attacked from the first combat unit created, using the SparCraft system to simulate whether or not the bot could a fight against known enemy units, retreating automatically if it thought it could not. This addition, combined with some additional bug fixes led to the victory. Skynet and Aiur both implemented strategy learning in 2013, which is evident from \ref{fig:aiide2013} in which both Skynet and Aiur's win percentage over time climb dramatically from the first few rounds of the tournament to the later rounds. Ximp unfortunately crashed all games on the Fortress map and lost 10\% of its games for free as a result, which meant it could have possibly came in 2nd place if not for all those crashes.

@@

\subsection{SSCAIT}\label{sec:SSCAIT}
Student StarCraft AI Tournament\footnote{\url{http://sscaitournament.com/}} (SSCAIT) started in 2011 at Comenius University in Bratislava, Slovakia and immediately became the competition with the highest numbers of participants.
There are three fundamental differences between SSCAIT and the remaining two conferences:
\begin{enumerate}
  \item SSCAIT is an online-only event. Unlike AIIDE or CIG, it is not co-located with a scientific conference or any other real-world event.
  \item There are two phases of SSCAIT each year: a {\em competitive} phase, lasting for up to three weeks and a {\em sandbox} phase which runs for approximately eleven months each year. In other words, SSCAIT is live at all times with only a few short interruptions.
  \item All the games are publicly {\em streamed live} 24 hours a day. The stream uses a custom observer script \cite{mattsson2015automatic} designed specifically to improve the viewer experience and creates a continuous feedback loop for the participants who can watch their bots play against the others.
\end{enumerate}

>>>>

In 2011 the University of Alberta hosted the competition, with organization by Michael Buro and
David Churchill\footnote{\url{https://skatgame.net/mburo/sc2011/}}. Due to a lack of entrants in tournament categories 1-3 in the 2010 competition, it was
decided that only the full game category would be played in the 2011 competition. Another important
change in the 2011 competition was the introduction of automated tournament-managing software running
StarCraft games simultaneously on 20 computers, allowing a total of 2340 games to be played in the five days that the tournament ran. 
This increase in games played also allowed the tournament
to switch to a round-robin format, eliminating the ``luck'' factor of the pairings inherent in bracket
style tournaments. The bot that achieved the highest win percentage over the course of the competition would
be determined the winner. Also, the competition became open-source, in an effort 
to prevent possible cheating and to promote healthy competition in future tournaments by giving
newcomers and easier entry point by basing their design off of previous bots.

@@

The University of Alberta also hosted the 2012 competition, with the major difference from the 2011
competition being the addition of file reading and writing for learning. Bots could now write information to disk during a
match, and then read the information during other matches, allowing them to adjust strategies based
on previous results. 6 of the 10 entrants used this feature to aid in strategy selection, including the
top 4 finishers. More improvements to the tournament environment also meant that a total of 8279 games
could now be played in the same time period. Results are shown in Table \ref{tab:aiide2012}.

@@

A total of 8 bots competed in the 2013 AIIDE competition with many of the same names from the 2012 competition, and some key updates were made to existing bots for 2013 which shook up the results from the previous years. A total of 5597 games were played during the 5 day tournament. UAlbertaBot took 1st place with a dominant 84.49\% win rate, with Skynet in 2nd with 72.77\%. The major addition to UAlbertaBot was a combat simulation package called SparCraft. UAlbertaBot was re-engineered so that it now always attacked from the first combat unit created, using the SparCraft system to simulate whether or not the bot could a fight against known enemy units, retreating automatically if it thought it could not. This addition, combined with some additional bug fixes led to the victory. Skynet and Aiur both implemented strategy learning in 2013, which is evident from \ref{fig:aiide2013} in which both Skynet and Aiur's win percentage over time climb dramatically from the first few rounds of the tournament to the later rounds. Ximp unfortunately crashed all games on the Fortress map and lost 10\% of its games for free as a result, which meant it could have possibly came in 2nd place if not for all those crashes.

@@

\subsection{SSCAIT}\label{sec:SSCAIT}
The Student StarCraft AI Tournament\footnote{\url{http://sscaitournament.com/}} (SSCAIT) started in 2011 at Comenius University in Bratislava, Slovakia and has been well known for being the Starcraft AI competition with the highest number of total participants. Started as a part of a course in artificial intellignece at Comenius University, each SSCAI tourament includes several dozen student submissions from this course in addition to submissions from across the globe. 
There are three fundamental differences between SSCAIT and the remaining two conferences:
\begin{enumerate}
  \item SSCAIT is an online-only event. Unlike AIIDE or CIG, it is not co-located with a scientific conference or any other real-world event.
  \item There are two phases of SSCAIT each year: a {\em competitive} phase, lasting for up to three weeks and a {\em sandbox} phase which runs for approximately eleven months each year. In other words, SSCAIT is live at all times with only a few short interruptions.
  \item All the games are publicly {\em streamed live} 24 hours a day. The stream uses a custom observer script \cite{mattsson2015automatic} designed specifically to improve the viewer experience and creates a continuous feedback loop for the participants who can watch their bots play against the others.
\end{enumerate}
